Using physical simulations to predict children's and adults' preference for physical interactions

### 2021-02-03 paper/paper2021.Rmd
model table with beta=1

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
```

@@

sh_subj_contain = split_half_subj_cor(subset(combo_dat, relation=="contain"))

sh_subj_support = split_half_subj_cor(subset(combo_dat, relation=="support"))

# split items in half, check cor between halves...GK: I don't think we care about this

@@

Finally, we examined the split-half reliability of adults' choices by repeatedly splitting the samples in half and testing the correlation between the halves.
Based on 100 random samples, the split-half reliability for adults is *r* = `r round(mean(sh_subj_adult), 2)` (sd=`r round(sd(sh_subj_adult), 2)`).

@@

## Results
\vspace{-4pt}
Were children more consistent in their choices on support or containment trials?
As for adults, averaging each child's binary responses for each trial type revealed a stronger preference for containment relations ($M=$ `r round(ch_rel_means[1,]$M,2)`) than support relations ($M=$ `r round(ch_rel_means[2,]$M,2)`; paired `r t_apa(ch_supp_cont, format="rmarkdown", print=F)`).
Overall, children did not show a significant preference for choosing the support relation. <!-- [add a regression with trial_type * age (3 - 18=adult)?] -->
Figure 4 shows the proportion of participants' choice of the designed target relation broken down by relation type and age group, showing a preference for the designed relations that increased with age, and that was stronger for containment.
We examined children's choice of target objects to determine if children were choosing randomly, or had consistent preferences for some objects. 
Table 1 shows the proportion of children ("child" column) choosing the designed target relation for each of the 20 trials, alongside the adult choice proportions from Experiment 1. 
Based on the binomial distribution, any of the trials on which more than 41 of the 66 participants agree (i.e., >0.64 or <0.36) significantly differ (*p*<.05) from chance. <!-- stricter crit for mult comparisons? Bonf=.0025 -->
Children significantly preferred the designed relation on eight of the 10 containment trials, while only significantly preferring two of the 10 support relation choices.
On the remaining two containment trials and the other eight support trials, children's preferences also did not significantly differ from chance against the designed relation.
The split-half reliability for children's choices is *r* = `r round(mean(sh_subj_child), 2)` (sd=`r round(sd(sh_subj_child), 2)`).

@@

The overall split-half reliability of children and adults on containment trials is *r* = `r round(mean(sh_subj_contain), 2)` (sd=`r round(sd(sh_subj_contain), 2)`), and on support trials is *r* = `r round(mean(sh_subj_support), 2)` (sd=`r round(sd(sh_subj_support), 2)`).
<!-- [should we separately report these for adults and children?] -->
```{r adults-vs-children, fig.env = "figure", fig.pos = "h", fig.align="center", fig.width=3.3, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Comparison of children's vs. adults' preferences on each trial (dot), with bootstrapped 95\\% confidence intervals."}
# add vert and horiz error bars 
combo_dat %>% group_by(drop, target, relation, ID) %>%
  group_by(drop, target, relation, AgeGroup) %>%
  tidyboot_mean(chose_target, na.rm=T) %>%
  select(-n, -empirical_stat) %>%
  pivot_wider(names_from = AgeGroup, values_from = c(mean, ci_lower, ci_upper)) %>%
  ggplot(aes(x=mean_adult, y=mean_child, group=relation, color=relation)) + 
    geom_point(alpha=.6) + 
    geom_errorbar(aes(ymin=ci_lower_child, ymax=ci_upper_child), alpha=.6) +
    geom_errorbarh(aes(xmin=ci_lower_adult, xmax=ci_upper_adult), alpha=.6) + 
    theme_bw() + ylim(0,1) + xlim(0,1) + 
    xlab("Adults' Choice Proportions") + ylab("Children's Choice Proportions") + 
    geom_abline(intercept=0, slope=1, linetype="dashed") +
    geom_smooth(method = "lm") +
    theme(legend.position = "bottom")
  #geom_text(aes(label=target)) #

@@

### Results 
The best-fitting feature was $Pr(support)$ for both children ($\beta =$ `r round(best_fits$beta[2],2)`, MSE = `r round(best_fits$mse[2], 3)`) and adults ($\beta =$ `r round(best_fits$beta[1],2)`, MSE = `r round(best_fits$mse[1], 3)`).\footnote{For $\beta=1$, adult MSE = `r round(supp_prob_beta1_cors$mse[1],3)`, child MSE = `r round(supp_prob_beta1_cors$mse[2],3)` for $Pr(support)$.}
<!-- other support-related features also fared well -->
Figure 6 shows this model's fitted preferences plotted against those of children and adults.
Although the model captured people's overall mean preference for the target relations, the trial-level correlation with human behavior was negative (adults $r=$ `r round(supp_prob_beta_opt_cors$r[1],2)`, children $r=$ `r round(supp_prob_beta_opt_cors$r[2],2)`, with similar values for $\beta = 1$).
The support response sharpness model fit slightly worse (children $\beta = 0.81$, MSE = 0.016) and adults ($\beta = 1.94$, MSE = 0.054), but its responses were positively correlated with people's (adults $r =$ `r round(supp_resp_sharp_beta_opt_cors$r[1], 2)`, children $r =$ `r round(supp_resp_sharp_beta_opt_cors$r[2],2)`).\footnote{For $\beta=1$, adult MSE = `r round(supp_resp_sharp_beta1_cors$mse[1],3)`, child MSE = `r round(supp_resp_sharp_beta1_cors$mse[2],3)` for support response sharpness, with very similar correlations.}

```{r human-vs-model-relational-choices, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.25, fig.height=4.75, set.cap.width=T, num.cols.cap=1, fig.cap = "Comparison of $Pr(support)$ model's vs. participants' choices for each type of trial, with 95\\% CIs."}
pos = position_dodge(width = .9)

>>>>

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
require(kableExtra)
```

@@

#sh_subj_contain = split_half_subj_cor(subset(combo_dat, relation=="contain"))
sh_ch_contain = split_half_subj_cor(subset(combo_dat, relation=="contain" & AgeGroup=="child"))
sh_ad_contain = split_half_subj_cor(subset(combo_dat, relation=="contain" & AgeGroup=="adult"))
#sh_subj_support = split_half_subj_cor(subset(combo_dat, relation=="support"))
sh_ch_support = split_half_subj_cor(subset(combo_dat, relation=="support" & AgeGroup=="child"))
sh_ad_support = split_half_subj_cor(subset(combo_dat, relation=="support" & AgeGroup=="adult"))
# split items in half, check cor between halves...GK: I don't think we care about this

@@

Finally, we examined the split-half reliability of adults' choices by repeatedly splitting the samples in half and testing the correlation between the halves.
Based on 100 random samples, the split-half reliability for adults was *r* = `r round(mean(sh_subj_adult), 2)` (sd=`r round(sd(sh_subj_adult), 2)`).
The split-half reliability of adults' choices on containment trials was *r* = `r round(mean(sh_ad_contain), 2)` (sd=`r round(sd(sh_ad_contain), 2)`), and on support trials was *r* = `r round(mean(sh_ad_support), 2)` (sd=`r round(sd(sh_ad_support), 2)`).

@@

## Results
\vspace{-4pt}
Were children more consistent in their choices on support or containment trials?
As for adults, averaging each child's binary responses for each trial type revealed a stronger preference for containment relations ($M=$ `r round(ch_rel_means[1,]$M,2)`) than support relations ($M=$ `r round(ch_rel_means[2,]$M,2)`; paired `r t_apa(ch_supp_cont, format="rmarkdown", print=F)`).
Overall, children did not show a significant preference for choosing the support relation. <!-- [add a regression with trial_type * age (3 - 18=adult)?] -->
Figure 4 shows the proportion of participants' choice of the designed target relation broken down by relation type and age group, showing a preference for the designed relations that increased with age, and that was stronger for containment.
We examined children's choice of target objects to determine if children were choosing randomly, or had consistent preferences for some objects. 
Table 1 shows the proportion of children ("child" column) choosing the designed target relation for each of the 20 trials, alongside the adult choice proportions from Experiment 1. 
Based on the binomial distribution, any of the trials on which more than 41 of the 66 participants agree (i.e., >0.64 or <0.36) significantly differ (*p*<.05) from chance. <!-- stricter crit for mult comparisons? Bonf=.0025 -->
Children significantly preferred the designed relation on eight of the 10 containment trials, while only significantly preferring two of the 10 support relation choices.
On the remaining two containment trials and the other eight support trials, children's preferences also did not significantly differ from chance against the designed relation.
The split-half reliability for children's choices is *r* = `r round(mean(sh_subj_child), 2)` (sd=`r round(sd(sh_subj_child), 2)`).
The split-half reliability of children's choices on containment trials was *r* = `r round(mean(sh_ch_contain), 2)` (sd=`r round(sd(sh_ch_contain), 2)`), and on support trials was *r* = `r round(mean(sh_ch_support), 2)` (sd=`r round(sd(sh_ch_support), 2)`).

@@

```{r adults-vs-children, fig.env = "figure", fig.pos = "h", fig.align="center", fig.width=3.3, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Comparison of children's vs. adults' preferences on each trial (dot), with bootstrapped 95\\% confidence intervals."}
# add vert and horiz error bars 
combo_dat %>% group_by(drop, target, relation, ID) %>%
  group_by(drop, target, relation, AgeGroup) %>%
  tidyboot_mean(chose_target, na.rm=T) %>%
  select(-n, -empirical_stat) %>%
  pivot_wider(names_from = AgeGroup, values_from = c(mean, ci_lower, ci_upper)) %>%
  ggplot(aes(x=mean_adult, y=mean_child, group=relation, color=relation)) + 
    geom_point(alpha=.6) + 
    geom_errorbar(aes(ymin=ci_lower_child, ymax=ci_upper_child), alpha=.6) +
    geom_errorbarh(aes(xmin=ci_lower_adult, xmax=ci_upper_adult), alpha=.6) + 
    theme_bw() + ylim(0,1) + xlim(0,1) + 
    xlab("Adults' Choice Proportions") + ylab("Children's Choice Proportions") + 
    geom_abline(intercept=0, slope=1, linetype="dashed") +
    geom_smooth(method = "lm") +
    theme(legend.position = "bottom")
  #geom_text(aes(label=target)) #

@@

### Results

```{r human-vs-model-table, results="asis"}
# now we just have a table of selected features that were moderately correlated with people's choices
load(here("models/model_preds_beta1.RData"))

described_features = c("support_response_sharpness_C=1",
                       "normed_velocity_std_after_first_collision_objects=target",
                       "normed_velocity_std_after_first_collision_objects=drop",
                       "avg_max_radius_objects=target",
                       "avg_max_radius_objects=drop",
                       "obj_final_position_invstd_objects=target",
                       "obj_final_position_invstd_objects=drop",
                       "avg_final_radius_objects=target",
                       "avg_final_radius_objects=drop",
                       "avg_len",
                       "len_std",
                       "support_probability")

supp_trials <- supp_trials %>% rename(`Support` = r) %>%
  select(-mse)

cont_trials <- cont_trials %>% rename(`Contain` = r) %>%
  select(-mse)

# combine all_trials, supp_trials, and cont_trials
mod_tab <- all_trials %>% select(-mse) %>%
  rename(`All` = r) %>% 
  left_join(supp_trials) %>%
  left_join(cont_trials) %>%
  filter(is.element(feature, described_features)) %>%
  arrange(desc(`All`), feature)  %>%
  mutate(Age = ifelse(Age=="children", "Child:", "Adult:")) %>%
  pivot_wider(names_from=Age, values_from=c(`All`, `Support`, `Contain`), 
              names_glue = "{Age} {.value}") %>%
  arrange(desc(`Adult: All`)) %>%
  mutate(feature = case_when(feature == 'support_response_sharpness_C=1' ~ 'Supp. Sharpness',
                             feature == 'support_probability' ~ 'Pr(support)',
                             feature == 'normed_velocity_std_after_first_collision_objects=target' ~ 'Target Velocity',
                             feature == 'normed_velocity_std_after_first_collision_objects=drop' ~ 'Drop Obj. Velocity',
                             feature == 'avg_len' ~ 'Avg. Distance',
                             feature == 'len_std' ~ 'Distance SD',
                             feature == 'obj_final_position_invstd_objects=target' ~ 'Target Pos. 1/SD',
                             feature == 'obj_final_position_invstd_objects=drop' ~ 'Drop Obj Pos. 1/SD',
                             feature == 'avg_final_radius_objects=target' ~ 'Avg. Target Radius',
                             feature == 'avg_final_radius_objects=drop' ~ 'Avg. Drop Obj Radius',
                             feature == 'avg_max_radius_objects=drop' ~ 'Max Drop Obj Radius',
                             feature == 'avg_max_radius_objects=target' ~ 'Max Target Radius',
                             TRUE ~ feature))

#modtab <- xtable::xtable(mod_tab, digits=2, caption = "Correlations of model and human responses.")
#modtab %>% kable(digits=2)
#print(modtab, type="latex", comment = F, include.rownames = F, table.placement = "H")

mod_tab <- as.matrix(mod_tab)
# bold max value per column
for(c in 2:ncol(mod_tab)) {
  mod_tab[,c] = round(as.numeric(mod_tab[,c]), 2)
  max_row = which(as.numeric(mod_tab[,c])==max(as.numeric(mod_tab[,c])))
  mod_tab[max_row,c] = cell_spec(mod_tab[max_row,c], "latex", bold = T)
}

kable(mod_tab, "latex", booktabs = T, escape=F, digits = 2, caption="Correlations ($r$) of model and human responses.") %>%
  column_spec(column = 2:ncol(mod_tab), width=c("1cm","1cm","1.3cm","1.3cm","1.3cm","1.3cm")) %>%
  kable_styling(latex_options = "scale_down") 
```

<!-- The best-fitting feature was $Pr(support)$ for both children ($\beta =$ `r round(best_fits$beta[2],2)`, MSE = `r round(best_fits$mse[2], 3)`) and adults ($\beta =$ `r round(best_fits$beta[1],2)`, MSE = `r round(best_fits$mse[1], 3)`).\footnote{For $\beta=1$, adult MSE = `r round(supp_prob_beta1_cors$mse[1],3)`, child MSE = `r round(supp_prob_beta1_cors$mse[2],3)` for $Pr(support)$.}
Figure 6 shows this model's fitted preferences plotted against those of children and adults.
Although the model captured people's overall mean preference for the target relations, the trial-level correlation with human behavior was negative (adults $r=$ `r round(supp_prob_beta_opt_cors$r[1],2)`, children $r=$ `r round(supp_prob_beta_opt_cors$r[2],2)`, with similar values for $\beta = 1$).
The support response sharpness model fit slightly worse (children $\beta = 0.81$, MSE = 0.016) and adults ($\beta = 1.94$, MSE = 0.054), but its responses were positively correlated with people's (adults $r =$ `r round(supp_resp_sharp_beta_opt_cors$r[1], 2)`, children $r =$ `r round(supp_resp_sharp_beta_opt_cors$r[2],2)`).\footnote{For $\beta=1$, adult MSE = `r round(supp_resp_sharp_beta1_cors$mse[1],3)`, child MSE = `r round(supp_resp_sharp_beta1_cors$mse[2],3)` for support response sharpness, with very similar correlations.}
-->

```{r human-vs-model-relational-choices, eval=F, include=F, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.25, fig.height=4.75, set.cap.width=T, num.cols.cap=1, fig.cap = "Comparison of $Pr(support)$ model's vs. participants' choices for each type of trial, with 95\\% CIs."}
pos = position_dodge(width = .9)
